Estimating dark matter halo concentration using the integrated mass profile

### 2015-11-11 paper/paper.tex
puliendo texto

We generate in total $400$ mock halos split into four different groups
of $100$ halos each.  The four groups differ in the total number of
particles for their halos: $20$, $200$, $2000$ and $20000$.  
Inside each group the halos have random concentration values in the range
$1<c<20$ with a uniform distribution.  
For all these halos we find  the concentration values using the
density, velocity and integrated mass methods described in the
previous section.

We quantify the difference between the expected $c_{in}$ and measured
$c_{out}$ values by 
%
\begin{equation}
D=(c_{in}-c_{out})/c_{in},
\label{eq:D}
\end{equation}
%
and
%
\begin{equation}
\avg{|D|}=\frac{1}{\left|{\cal{H}}\right|}\sum_{{\cal{H}}} |D|,
\end{equation}
%
where ${\cal{H}}$ corresponds to a set of mock halos, and
$\left|{\cal{H}}\right|$ is the number of haloes in ${\cal{H}}$. 
  The larger the value of $\avg{|D|}$, the larger the discrepace of
  the concentrations obtained for that method.

@@

Figure \ref{fig:error} shows the behaviour of $\avg{|D|}$ as a function of
halo particle number for the three different methods to estimate the
concentration.

At fixed particle numbers our method always shows the lowest
$\avg{|D|}$ values compared to the other two methods.  Its accuracy is
on the order of $10\%$ for $20$ particles in the halo, going down to
$0.01\%$ for halos with $20000$ particles.  The dependence of
$\avg{|D|}$ with the particle number $N$ goes approximately as
$\avg{|D|}\propto N^{-1/2}$, hinting that the accuracy of the method
is related to a decrease of Poisson noise.

The method based on the maximum of the circular velocity shows a
similar behaviour $\avg{|D|}\propto N^{-1/2}$.  Its accuracy is $2-5$
times less than in our method, on the order of $20\%$ for $20$
particle halos and $0.5\%$ for $20000$ particle halos. {\bf The method
  based on the direct density fit shows the lowest accuracy for a low
  particle number. As it is discussed in \citep{Munoz2011}, density
  binning for halos with particle numbers below $\sim200$ leads to a
  biased estimation of the mass density profile, and therefore to a
  biased estimation of the concentration parameter. This behaviour is
  evident in the large values of $\avg{|D|}$ for halos with number of
  particles below ~200 and an intermediate accuracy between the other
  two methods for a high particle number.}

@@

{\bf The results presented in the previous section show that in an
  idealized setup (perfect spherical halos, pure NFW desity profiles,
  total isolation) our method of the mass profile performs better than
  the other two methods commonly used to measure concentrations in
  simulated dark matter haos. However, actual simulated dark matter
  halos are much more complex. They are not isolated, many of them
  experience mergers that take them out of the equilibrium. Halos may
  have also plenty of substructure, etc. All these effects disturb the
  mass density profile, taking it away from the ideal NFW profile we
  assume in this wok. In order to see the real performance of our new
  method use it to estimate the concentration parameter in dark matter
  halos obtained from a cosmological Nbody simulation and compare with
  the other two calsical methods.}

We use data from the MultiDark cosmological simulation that follows
the non-linear evolution of a dark matter density field sampled with
$2048^3$ particles over a cubic box of $1000\ \hMpc$ on a side.  The
data is publicly available through \url{http://www.cosmosim.org/}.
More details about the structure of the database and the simulation
can be found in \citep{2013AN....334..691R}.

We build a sample of all halos located in a cubic sub-volume of $100$
\hMpc on a side centered on the most massive halo in the simulation at
$z=0$ which corresponds to the \texttt{miniMDR1} tables in the
database.  From this sample we select all the halos at $z=0$ detected
with a Friends-of-Friends (FoF) algorithm with masses in the interval
$10^{12}\leq M_{\rm FoF}/\hMsun \leq 10^{15}$.  The FoF algorithm used
with a linking length of $0.17$ times the {\bf mean} inter-particle
distance. This choice translates into an overdensity $\Delta_h\sim
400-700$ dependent on the halo concentration \citep{More2011}.
Finally, we select from the database all the particles that belong to
each FoF group.

>>>>

We generate in total $400$ mock halos split into four different groups
of $100$ halos each.  
The four groups differ in the total number of particles for their
halos: $20$, $200$, $2000$ and $20000$.   
Inside each group the halos have random concentration values in the range
$1<c<20$ with a uniform distribution.  
We find the concentration values for all these halos using the
density, velocity and integrated mass methods described in the
previous section.

We quantify the difference between the expected $c_{in}$ and measured
$c_{out}$ values by 
%
\begin{equation}
D=(c_{in}-c_{out})/c_{in},
\label{eq:D}
\end{equation}
%
and
%
\begin{equation}
\avg{|D|}=\frac{1}{\left|{\cal{H}}\right|}\sum_{{\cal{H}}} |D|,
\end{equation}
%
where ${\cal{H}}$ corresponds to a set of mock halos, and
$\left|{\cal{H}}\right|$ is the number of haloes in ${\cal{H}}$. 
A large discrepacy between the estimated and input concentrations are
quantified by a large $\avg{|D|}$.

@@

Figure \ref{fig:error} shows the behaviour of $\avg{|D|}$ as a function of
halo particle number for the three different methods.

At fixed particle numbers our method always shows the lowest
$\avg{|D|}$ values compared to the other two.
Its accuracy is on the order of $10\%$ for $20$ particles in the halo,
decreasing to $0.01\%$ for halos with $20000$ particles.  
The dependence of $\avg{|D|}$ with the particle number $N$
goesapproximately as $\avg{|D|}\propto N^{-1/2}$.   
This hints that the accuracy of the method could be due to a decrease
of Poisson noise.

The method based on the maximum of the circular velocity shows a
similar behaviour $\avg{|D|}\propto N^{-1/2}$.  
Its accuracy is $2-5$ times lower than in our method, on the order of
$20\%$ for $20$  particle halos and $0.5\%$ for $20000$ particle
halos. 
The method  based on the direct density fit shows the lowest accuracy for a low
  particle number. 
  As it is discussed in \citep{Munoz2011}, density binning for halos
  with particle numbers below $\sim200$ leads to a biased estimation
  of the mass density profile, and therefore to a biased estimation of
  the concentration parameter. 
This behaviour is evident in the large values of $\avg{|D|}$ for halos
with number of particles below ~200 and an intermediate accuracy
between the other two methods for a high particle number.

@@

The results presented in the previous section show that in an
idealized setup (pure NFW desity profiles, perfect spherical halos in
total isolation) our method performs better than the other two
methods commonly used in the literature.

However, simulated dark matter halos in a cosmological context are
complex. 
They are not isolated, many of them  experience mergers that take them
out of the equilibrium. 
Halos may have also plenty of substructure, etc. All these effects
disturb the mass density profile, taking it  away from the ideal NFW
profile. 
In order to see the completeley explore performance of our method, we use
it to estimate the  concentration parameter in dark matter halos
obtained from a cosmological N-body simulation and compare with the
other two classical methods.

We use data from the MultiDark cosmological simulation that follows
the non-linear evolution of a dark matter density field sampled with
$2048^3$ particles over a cubic box of $1000\ \hMpc$ on a side.  
The data is publicly available at \url{http://www.cosmosim.org/}.
More details about the structure of the database and the simulation
can be found in \citep{2013AN....334..691R}.

%%%%%%%

We use a halo sample containing all the halos located in a cubic
sub-volume of $100$ \hMpc on a side centered on the most massive halo
in the simulation at $z=0$  which corresponds to the \texttt{miniMDR1}
table in the  database.  
From this sample we select all the halos at $z=0$ detected
with a Friends-of-Friends (FoF) algorithm with masses in the interval
$10^{12}\leq M_{\rm FoF}/\hMsun \leq 10^{14}$.  
The FoF algorithm used a linking length of $0.17$ times the mean
inter-particle distance. This choice translates into an overdensity
$\Delta_h\sim  400-700$ dependent on the halo concentration \citep{More2011}.
Finally, we select from the database all the particles that belong to
each FoF group.
